---
title: "Homework 4"
author: "Benise Limon"
format: docx
editor: visual
execute: 
  message: false
  warning: false
---

```{r}
#call the packages I want to use 
library(tidyverse)
library(here)
library(lterdatasampler)
library(performance)
library(broom)
library(ggeffects)
library(car)
library(naniar)
library(flextable)
```

How does fish length predict fish weight for trout perch (across all sample years)?\
This data set, like most observational environmental or ecological data sets, is noisy (i.e. a lot of spread between points). It is also somewhat messy. You are expected to read the metadata, protocols, and all relevant information on the EDI portal before working with the data.

# Problem 1

1\. Write your null and alternative hypotheses in mathematical and biological terms.

Biological:

H0: Fish length does not predict fish weight for trout perch across all sample years.

HA: Fish length does predict fish weight for trout perch across all sample years.

Mathematical:

H0: β1 = 0

HA: β1 ≠ 0

\
2. Create a visualization of the missing data for the filtered data set containing the\
observations you will use.

```{r read-data}
#make an object that reads the CSV file containing the data 
fish <- read.csv(here("data/ntl6_v12.csv"))
```

```{r filtered-data}
#filter the data to only include trout 
fish_clean <- fish %>% 
  filter(spname == "TROUTPERCH")
```

```{r missing-data-vis}
#visualize missing data 
gg_miss_var(fish_clean) +
  #add a caption of how missing data affects hypothesis 
  labs (caption = str_wrap("Figure 1. There are 200 missing observations for fish weight. This affects our hypothesis to understand if fish length predicts fish weight.")) +
  theme(
    plot.caption = element_text(hjust = 0)
  )
```

\
a. Write an accompanying caption explaining how/if the missing data is relevant to\
your hypotheses.

There are 200 observations from the "weight" column are missing. Since it is missing, we will have to delete the null values from the data set, meaning there are less observations to work with. The linear regression is intended to see if fish length can predict fish weight, missing some values of fish weight gives us less data to base our analysis on.

\
3. Run your test.

```{r model-object}
#create a model object containing the linear model for if fish length predicts weight using the filtered data frame
modelobject <- lm(weight ~ length, data = fish_clean)
```

\
4. Visually check the assumptions of your test. Display your diagnostic plots in a grid.

```{r display-model}
# makes the viewer pane show a 2x2 grid of plots
# format: par(mfrow = c(number of rows, number of columns))
par(mfrow = c(2, 2))
plot(modelobject)
```

```{r dev-off}
# turns off the 2x2 grid - pop this under the code chunk where you set the 2x2 grid
dev.off()
```

\
5. For each diagnostic plot, describe in 1-2 sentences what it is showing you, and what you\
decide after looking at the plot.

Residuals vs Fitted:

Scale-location:

Normal QQ:

Residuals vs Leverage:

\
6. Display the results from summary() using your model object.

```{r summary-table}
# store the model summary as an object
model_summary <- summary(modelobject)

model_summary
```

7\. Create a table that summarizes the ANOVA table. Make sure this table has informative\
column and row names and p-values displayed without scientific notation.

```{r anova-table}
# store the ANOVA table as an object
# anova(): special function to get analysis of variance tables for a model
model_squares <- anova(modelobject)

model_squares
```

```{r}
# don't name this chunk! some intricacies with Quarto: do not name chunks with tables in them

model_squares_table <- tidy(model_squares) %>% 
  # round the sum of squares and mean squares columns to have 5 digits (could be less)
  mutate(across(sumsq:meansq, ~ round(.x, digits = 2))) %>% 
  # round the F-statistic to have 1 digit
  mutate(statistic = round(statistic, digits = 1)) %>% 
  # replace the very very very small p value with < 0.001
  mutate(p.value = case_when(
    p.value < 0.001 ~ "< 0.001"
  )) %>% 
  # rename the length cell to be meaningful
  mutate(term = case_when(
    term == "length" ~ "Length (mm)",
    TRUE ~ term
  )) %>% 
  # make the data frame a flextable object
  flextable() %>% 
  # change the header labels to be meaningful
  set_header_labels(df = "Degrees of Freedom", 
                    sumsq = "Sum of squares",
                    meansq = "Mean squares",
                    statistic = "F-statistic",
                    p.value = "p-value")

model_squares_table
```

\
8. In 1-2 sentences, describe how the ANOVA table relates to the information you get from\
the summary() object.

The ANOVA table relates to the information from the summary() object because the it is created using the anova function which extracts summary information from the model object, including the degrees of freedom, sum of squares, mean squares, and the p-value. The ANOVA table uses the summary of the model object to understand if there is a significant relationship between the predictor variable and what you are trying to predict.

\
9. In 2-3 sentences, summarize your results in prose with in-text references to test results. Include all relevant information.

The results from this linear regression were that fish length can be used to predict fish weight. The null hypothesis was rejected because the p-value was \<0.001. The high R\^2 value explains 81.58% of the variance all fish weight observations can be predicted by the model. The low mean squared value of 1.12 supports the alternative hypothesis because on average the model only differs from the actual values by 1.12g. The scatter plot visually shows the observations are clustered along the regression line, indicating the regression line is a good fit. Analysis of variance, F(1,288) = 1280 p\< 0.001, a = 0.05.

\
10. Create a visualization with model predictions and confidence intervals on top of the\
underlying data. Finalize your plot.

```{r model-predictions}
# extract model predictions using ggpredict
predictions <- ggpredict(modelobject, terms = "length")

predictions
```

```{r}
plot_predictions <- ggplot(data = fish_clean, 
                           aes(x = length, y = weight)) +
  # first plot the underlying data from fish_clean
  geom_point() +
  # then plot the predictions
  geom_line(data = predictions, 
            aes(x = x, y = predicted), 
            color = "green", linewidth = 1) +
  # then plot the 95% confidence interval from ggpredict
  geom_ribbon(data = predictions, 
              aes(x = x, y = predicted, ymin = conf.low, ymax = conf.high), 
              alpha = 0.2) +
  # theme and meaningful labels
  theme_bw() +
  labs(x = "Trout Perch length (mm)",
       y = "Trout Perch weight (g)", 
       caption = "Figure 2. A visualization of using fish length to predict fish weight in Trout Perch. A ___ increase in length (mm) predicts a ___ increase in weight (g)") +
        theme(
    plot.caption = element_text(hjust = 1))

plot_predictions
```

\
a. Write an accompanying caption.

A visualization of using fish length to predict fish weight in Trout Perch. A \_\_\_ increase in length (mm) predicts a \_\_\_ increase in weight (g)\

Checklist\
Your submission should include\
- A link to your public GitHub repo for this homework assignment that includes\
- An informative readme\
- File organization\
- At least 5 commits\
- Only relevant code and outputs, with no messages or warnings\
- Annotated, spaced out code\
- Written captions and results summaries\
- Additionally, your submission should be submitted to Canvas by 11:59 PM on Tuesday 23 May
